---
title: "MiniProject4"
format: html
editor: visual
---

# Question: 

Are some communities in Minneapolis more dangerous than other Communities in terms of crime? Are some police precincts more violent than others?

```{r echo=TRUE,message=FALSE}
# Look at day22(redlining) for leaftlet examples
library(tidyverse)
library(stringr)
library(tidyverse)
library(sf)     # R wrapper around GDAL/OGR
library(leaflet)    # for fortifying shapefiles
library(leaflet.extras)   # For leaflet heatmaps
```

## Read in shapefiles.

```{r echo=TRUE,message=FALSE}
# Read in the communities data
# DSN - location of shp files
# Layer - name of all the files in DSN (different extensions, same names)
mn_comm <- st_read(dsn="Communities-SHP/", layer="Minneapolis_Communities",
                   # Stop printing of reading in st_read
                   quiet=TRUE)
minn_pre <- st_read(dsn="Minneapolis_Police_Precincts-SHP/",
                         layer="Police_Precincts",
                    quiet=TRUE)
police_shootings <- st_read(dsn="Police_Officer_Involved_Shootings/",
                            layer="Police_Officer_Involved_Shootings",
                            quiet=TRUE)
police_force <- st_read(dsn="Police_Use_of_Force",
                        layer="Police_Use_of_Force",
                        quiet=TRUE)
# https://cms9files.revize.com/richfieldmn/Document_Center/Department/Community%20Development/Planning%20and%20Zoning/Cedar%20Avenue%20Corridor/CedarCorridorComprehensive.pdf
# https://www.dot.state.mn.us/corridorsofcommerce/about.html
# https://opendata.minneapolismn.gov/datasets/cityoflakes::goods-and-services-corridor/
corridors <- st_read(dsn="Goods_and_Services_Corridor",
                     layer="Goods_and_Services_Corridor",
                     quiet=TRUE)

# -93.029651 CenterLong not in minnneapolis
# 45.196298 CenterLat
police_shootings <- police_shootings |>
  filter(CenterLati != 45.196298)

# glimpse(mn_comm)
# glimpse(minn_pre)
# glimpse(police_shootings)
# glimpse(police_force)
```

## Build a base plot

```{r}
# Label fatal shootings as red, non-fatal as red.
fatalPal <- colorFactor(palette=c("orange", "red"),
                          domain=police_shootings$SubjectFat)
#glimpse(corridors)
pre <-
  leaflet() |> 
  addProviderTiles(providers$CartoDB.DarkMatter, group = "Dark") |>
  addTiles(group = "Default") |>
  addProviderTiles(providers$CartoDB.Positron, group = "Default2") |>
  addPolylines(data=corridors,
              weight=1,
              color="white") |>
  addCircleMarkers(data=police_shootings,
                   ~CenterLong,
                   ~CenterLati,
                   color=fatalPal(police_shootings$SubjectFat),
                   radius=4,
                   # popup=~as.character(SubjectFat),
                   stroke=FALSE, fillOpacity=1) |>
  addLegend(
    data = police_shootings,
    title = lapply("Police Brutality<br>(Fatal Shootings)",htmltools::HTML),
    pal = fatalPal,
    values = ~police_shootings$SubjectFat,
    opacity = 0.6,
    group = "Brut. Leg."
  ) |>
  addLayersControl(
    baseGroups = c("Dark", "Default", "Default2"),
    overlayGroups = c("Brut. Leg."),
    # options = layersControlOptions(collapsed = FALSE),
    position = "topleft"
  )
pre
```

## Scraping html

Scrape information from url containing neighborhoods by community. Couldn't find an open dataset of neighborhoods grouped by community. I could also just use a heatmap for neighborhoods to show more information.

```{r}
library(rvest)
url <- "https://homesmsp.com/2011/09/minneapolis-neighborhoods-by-community.html"
# Scrape url for strong elements (communities)
html_comm <- url |>
  read_html() |>
  html_elements("strong") |>
  html_text2()
#html_comm

# replace first param (exact string) with nothing for every string in list
comm <- gsub(" Community Neighborhoods – average price","",html_comm)
comm <- gsub("CALHOUN/ISLES", "Calhoun Isle", comm)
comm <- gsub("[/]", "", comm)
comm <- comm[c(-12,-13)] # remove info that are not communitys
comm <- str_to_lower(comm) # lower all casing
comm <- str_to_title(comm) # use title casing for communities
#comm

html_neigh_uls <- url |>
  read_html() |>
  # Main Content inside div with class=infinite-single-article-content
  # Use chrome dev tools
  html_elements(".infinite-single-article-content ul")
#html_neigh_uls

neigh_list <- list()
# loop through ul's in html_neigh
#html_neigh_uls

for (ul in html_neigh_uls){
  li_list <- ul |>
    html_elements("li") |>
    html_text2()
  # append list to list (each index is a neighborhood)
  neigh_list <- append(neigh_list, list(li_list))
}
#neigh_list

cleaned <- list()
for (list in neigh_list){
  # Find dots, $, 0-9 and comma and replace with nothing
  list <- gsub("[.$0-9,]", "", list)
  # Replace dashes/forward slah in neighborhood names with spaces instead
  list <- gsub("[-/]", " ", list)
  # Remove trailing white space from strings in list
  list <- trimws(list)
  #print(list)
  cleaned <- append(cleaned, list(list))
}
cleaned <- cleaned[-12] # remove anchor tag text from html
#cleaned

# Check that lists are same length list of list = length of comm. list.
#print(length(cleaned))
#print(length(comm))

#df <- data.frame(comm, cleaned)
#df
# Error in (function (..., row.names = NULL, check.rows = FALSE, check.names = TRUE, :
# arguments imply differing number of rows: 7, 13, 5, 6, 9, 2, 8, 11 (What does this mean in r?)

df1 <- data.frame()
for (x in 1:11){
  df <- data.frame(comm[[x]], cleaned[[x]])
  df1 <- rbind(df1, df)
}
#df1

# Change column names.
colnames(df1)[1] <- "Community"
colnames(df1)[2] <- "Neighborhood"
#df1

# Correct some differences in naming between scraped and open dataset.
police_force <- police_force |>
  mutate(Neighborho = gsub(" - ", " ", Neighborho),
         Neighborho = gsub("[.']", "", Neighborho))
# Check neighborhood names in police_force/new dataframe
filtered <- police_force |>
  distinct(police_force$Neighborho)

#view(filtered)
#view(df1)

# check non-matches and fix
df1 |>
  #left_join(filtered, by=c("Neighborhood"="police_force$Neighborho"))
  anti_join(filtered, by=c("Neighborhood"="police_force$Neighborho"))

# Add the missing rows. (Old = duplicates)
df1[nrow(df1) + 1,] <- c("Phillips", "Phillips West")
df1[nrow(df1) + 1,] <- c("Phillips", "East Phillips")
# Spacing might be off on this one, looks same.
df1[nrow(df1) + 1,] <- c("University", "Nicollet Island East Bank")
df1[nrow(df1) + 1,] <- c("Calhoun Isle", "ECCO")
df1[nrow(df1) + 1,] <- c("University", "Prospect Park East River Road")

#write.csv(df1, "some_name.csv")
```
## Aggregating data

Aggregate number of crimes and summary of information to attach to communitys in Minneapolis.

```{r}
# Join police_force dataset and 
# scraped neighbordhoods by community df.
# https://opendata.minneapolismn.gov/datasets/cityoflakes::police-use-of-force/about

# Filter down dataset.
columnsOfInterest <- c("SubjectInj",
                       "Neighborho",
                       "Race",
                       "Sex",
                       "ForceType",
                       "EventAge",
                       "Precinct")

aggregated <- police_force |>
  select(all_of(columnsOfInterest)) |>
    mutate(isMale = ifelse(Sex == "Male", 1, 0),
         isFemale = ifelse(Sex == "Female", 1, 0),
         isInjured = ifelse(SubjectInj == "Yes", 1, 0),
         # Take top 3 force used by police, as shown in their bar charts.
         isBodilyForce = ifelse(ForceType == "Bodily Force", 1, 0),
         isTaser = ifelse(ForceType == "Taser", 1, 0),
         isChemicalIrritant = ifelse(ForceType == "Chemical Irritant", 1, 0),
         isPoc = ifelse(Race == "White", 0, 1))
  
#glimpse(aggregated)

aggreg <- aggregated |>
  inner_join(df1, c("Neighborho" = "Neighborhood"))

#aggregated

summ <- aggreg |>
  group_by(Community) |>
  summarize(num_crimes = n(),
            # Remove na's, else you get NA as result.
            rateM = sum(isMale, na.rm=TRUE)/num_crimes * 100,
            rateF = sum(isFemale, na.rm=TRUE)/num_crimes * 100,
            rateOther = (1 - ((rateM/100) + (rateF/100))) * 100,
            rateInjured = sum(isInjured, na.rm=TRUE)/num_crimes * 100,
            rateBodilyForce = sum(isBodilyForce, na.rm=TRUE)/num_crimes * 100,
            rateTasered = sum(isTaser, na.rm=TRUE)/num_crimes * 100,
            rateChemicalIrritant = sum(isChemicalIrritant, na.rm=TRUE)/num_crimes * 100,
            ratePoc = sum(isPoc, na.rm=TRUE)/num_crimes * 100)
#view(summ)

# have to drop geometry to join with other spatial data frame (mn_comm)
summ <- st_drop_geometry(summ)
mn_comm <- mn_comm |>
  inner_join(summ, by=c("CommName"="Community"))
#mn_comm
```
## Thoughts

In the following choropleth map, community boundaries are drawn on the map. From the choropleth, we can conclude that some communities are more dangerous than others, but this could also be because the population size in the city is denser than in the suburbs. We can also see that you are more likely to be shot by an officer or killed in the more dangerous communities. Population size is a question that I would delve into in the future by finding a dataset for populations per neighborhood or by the community to further illustrate in this map.

```{r}
# https://rstudio.github.io/leaflet/choropleths.html

# Use continuous color pallete for chloropeth.
neighPal <- colorFactor(palette="YlOrRd",
                                 domain=mn_comm$num_crimes)

# Make format for community labels.
community <- "<strong>%s</strong>"
totalCrimes <- "Total Crimes: %d"
rateRace <- "Person of color: %.2f%%"
rateM <- "Male: %.2f%%"
rateF <- "Female: %.2f%%"
rateO <- "Other: %.2f%%"
rateInj <- "Suspect Injured: %.2f%%"
rateBody <- "Bodily Force Used: %.2f%%"
rateTaser <- "Tasered: %.2f%%"
rateChemical <- "Chemical Force: %.2f%%"

# Concatenate with <br> for newline in label.
labelFormat <- paste(community,
                   totalCrimes,
                   rateRace,
                   rateM,
                   rateF,
                   rateO,
                   rateInj,
                   rateBody,
                   rateTaser,
                   rateChemical,
                   sep="<br>")
labels <- sprintf(
  labelFormat,
  mn_comm$CommName, 
  mn_comm$num_crimes,
  mn_comm$ratePoc,
  mn_comm$rateM, 
  mn_comm$rateF,
  mn_comm$rateOther,
  mn_comm$rateInjured,
  mn_comm$rateBodilyForce,
  mn_comm$rateTasered,
  mn_comm$rateChemicalIrritant) |> 
  lapply(htmltools::HTML)

# Plot the aggregated community data.
mn <- pre |>
  addPolygons(data=mn_comm,
              weight = 1,
              color = "white",
              fillColor = ~neighPal(num_crimes),
              fillOpacity = 0.4,
              highlight = highlightOptions(weight=3,
                                           color="red",
                                           bringToFront=TRUE),
              label=labels,
              labelOptions = labelOptions(style=list(
                # Add css styling here.
                "font-weight"="normal", # step below bold
                "padding"="3px 8px",
                # rgb(255,255,255) = white, opacity
                # setting opacity doesnt work for some reason.
                "background"="rgba(255, 255, 255, .70)"),
                textsize = "15px",
                direction = "auto")) |>
  addLegend(
    data = mn_comm,
    title = "Total Crimes",
    pal = neighPal,
    values = ~num_crimes,
    opacity = 0.6,
    group = "Comm. Leg."
  ) |>
  # https://rstudio.github.io/leaflet/showhide.html
  addLayersControl(
    baseGroups = c("Dark", "Default", "Default2"),
    overlayGroups = c("Comm. Leg.", "Brut. Leg."),
    #options = layersControlOptions(collapsed = FALSE)
    position = "topleft"
  )
mn
```
## Are some police precincts more violent than others?

One of the most notable here is Precinct #1, which has used chemical force in 24% of the total crimes in their area. Meanwhile, other police precincts have stayed below 8% (next highest). Another notable difference is that the darker regions (more crime) have had more injuries sustained to the suspect, specifically precincts 1, 3, and 4. Precinct #1, being an outlier from the 3, was it the use of chemical force that lowered overall injuries (possibly incapacitating the suspect before getting any more violent?). As I analyzed the graphic, I noticed that many fatal police shootings were nearing a precinct's border and heading to the 'safer' neighborhoods. I thought the suspects could have been fleeing on main roads where travel was easier. To confirm this, I downloaded an open dataset from Minneapolis, added the good and services corridors that make travel easier, and confirmed my suspicion. Most of the shootings tend to happen around these roads. For example, in precinct 5, three shootings happened on Franklin Ave., which does not seem like a coincidence. I believe these suspects were fleeing from precinct 3 (most likely) or 4. Overall I think some precincts are more violent than others, but I cannot conclude this based on my observations. This could be a factor of more crime because the population is higher or because with bigger police forces (and less budget), bad recruits get in and push the numbers up.

```{r}
# Convert to doubles, join on Precinct
minn_pre <- minn_pre |>
  mutate("Precinct" = as.double(PRECINCT))
aggregated <- aggregated |>
  mutate("Precinct" = as.double(Precinct))

#minn_pre <- st_drop_geometry(minn_pre)
#glimpse(minn_pre)
#glimpse(police_force)
#glimpse(aggregated)

# Drop aggregated geometry column.
aggregated <- st_drop_geometry(aggregated)
# Aggregate data by precinct.
precint_data <- aggregated |>
  group_by(Precinct) |>
  summarize(num_crimes = n(),
            # Remove na's, else you get NA as result.
            rateM = sum(isMale, na.rm=TRUE)/num_crimes * 100,
            rateF = sum(isFemale, na.rm=TRUE)/num_crimes * 100,
            rateOther = (1 - ((rateM/100) + (rateF/100))) * 100,
            rateInjured = sum(isInjured, na.rm=TRUE)/num_crimes * 100,
            rateBodilyForce = sum(isBodilyForce, na.rm=TRUE)/num_crimes * 100,
            rateTasered = sum(isTaser, na.rm=TRUE)/num_crimes * 100,
            rateChemicalIrritant = sum(isChemicalIrritant, na.rm=TRUE)/num_crimes * 100,
            ratePoc = sum(isPoc, na.rm=TRUE)/num_crimes * 100)

#glimpse(precint_data)

# Join precint data.
precint_data <- precint_data |>
  inner_join(minn_pre, "Precinct") |>
  group_by(Precinct)

#glimpse(precint_data)

# https://stackoverflow.com/questions/68728154/leaflet-map-error-in-polygondata-defaultdata-dont-know-how-to-get-path-dat
precint_data <- sf::st_as_sf(precint_data)

# Make new labels for precinct data.
policeLabels <- sprintf(
  labelFormat,
  precint_data$Precinct, 
  precint_data$num_crimes,
  precint_data$ratePoc,
  precint_data$rateM, 
  precint_data$rateF,
  precint_data$rateOther,
  precint_data$rateInjured,
  precint_data$rateBodilyForce,
  precint_data$rateTasered,
  precint_data$rateChemicalIrritant) |> 
  lapply(htmltools::HTML)

# Create pallete for precinct dataframe.
policePal <- colorFactor(palette="Blues",
                                 domain=precint_data$num_crimes)
pre_pal <- colorFactor(palette="Set2",
                       domain=minn_pre$PRECINCT)
# Plot data.
pre |>
  addPolygons(data=precint_data,
              weight = 3,
              color = ~pre_pal(PRECINCT),
              dashArray = "3",
              fillColor = ~policePal(num_crimes),
              fillOpacity = 0.3,
              highlight = highlightOptions(weight=5,
                                           color="black",
                                           bringToFront=TRUE),
              label=policeLabels,
              labelOptions = labelOptions(style=list(
                "font-weight"="normal",
                "padding"="3px 8px",
                "background"="rgba(255, 255, 255, .70)"),
                textsize = "15px",
                direction = "auto")) |>
  addLegend(data = minn_pre,
            title = "Precincts",
            pal = pre_pal,
            values = ~PRECINCT,
            opacity = 0.6,
            group = "Prec. Leg.") |>
  addLegend(data = precint_data,
            title = "Total Crimes",
            pal= policePal,
            value = ~num_crimes,
            opacity = 0.6,
            group = "Crimes Leg.") |>
  # https://rstudio.github.io/leaflet/showhide.html
  addLayersControl(
    baseGroups = c("Dark", "Default", "Default2"),
    overlayGroups = c("Prec. Leg.", "Crimes Leg.", "Brut. Leg."),
    #options = layersControlOptions(collapsed = FALSE)
    position = "topleft"
  )
```




